---
output:
  html_notebook: default
  html_document: default
---






## Required packages 


```{r}
library(tidyr)
library(readr)
library(stringr)
library(dplyr)
library(Hmisc)
library(outliers)
library(InformationValue)
library(ggplot2)
library(corrplot)
library(rpart)
library(Metrics)
library(car)

```

## Executive Summary 


## Data 

The dataset downloaded from the UCI Machine Learning Repository is related to red variant of the Portuguese "Vinho Verde" wine.


```{r}
#read data
redwine <- read.csv("/Users/Vidya/Downloads/winequality-red.csv", sep = ";")

describe(redwine)
str(redwine)
summary(redwine)
table(redwine$quality)

ggplot(data = redwine, aes(redwine$quality)) + geom_histogram(binwidth = 1, bins = 10, color = "black", fill = "Light Blue")

```




The wines are given a score between 1 and 10, 1 being the lowest and 10 being the highest rating for a wine. 
Filter quality to see the score distribution for wines with quality rating of 8 or above. Factor was 

```{r}
best_wines <- redwine %>% filter(quality > 7)

```

There are no null values in the dataset. A correlaton matrix was generated to check whether any significant correlations existed between the features in the dataset. In the generated correlations heatmap, negative correlations are in blue and positive ones in red color.

```{r}
summary(best_wines)
corr_redwine <- cor(redwine)
corrplot(corr_redwine, method = "number", number.font = 2)


```

There is moderate correlation between citric acid and volatile acidity, density and alcohol, quality and volatile acidity and quality and alcohol.
The boxplot for the volatile acidity shows that red wine with lower volatile acidity have a higher quality rating. The distribution of the percent alcohol content of the wine showed a higher median readings in wines rated 6 or higher. According to the visualisations, red wine with higher alcohol content and lower volatile acidity levels were rated higher in quality.


```{r}
ggplot(data = redwine, aes(x = density, y = alcohol))+ geom_point(color = "navy blue")+ geom_smooth(method = "lm")
ggplot(data = redwine, aes(x = citric.acid, y = volatile.acidity))+ geom_point(color = "navy blue")+ geom_smooth(method = "lm")
ggplot(redwine, aes (y=volatile.acidity, x=quality, group = quality)) + geom_boxplot( color = "light blue")

ggplot(data = redwine, aes(x = quality, y = alcohol, group = quality)) + geom_boxplot( color = "black", fill ="orange")+ scale_x_continuous(breaks = 0:10)
```
The wine was labelled low, average or high based on the quality score. Wine scoring 3 or lower than 3 were labeled Low, wines with a score ranging from 4 to 7 were labeled as Average. Wines that scored 8 or above were labeled high.

```{r}


#redwine <- mutate(redwine, wine_quality = ifelse(quality %in% 1:4, "Low", ifelse(quality %in% 5:7, "Average", "High")))
#table(redwine$wine_quality)

```
 
The data was split into test and training sets

```{r}
set.seed(123)

n <- nrow(redwine)
n_train <- round(0.75*n)

train_indices <- sample(1:n, n_train)
wine_train <- redwine[train_indices ,]
wine_test <- redwine[-train_indices ,]
dim(wine_train)
dim(wine_test)


```




Train a gini-based model

```{r}
wine_tree <- rpart(formula = quality~.,
                   data = wine_train,
                   method = "class",
                   parms = list(split = "gini"))


```

Plot the decison tree

```{r}

wine_tree
rpart.plot(wine_tree)
```





# Generate predictions on the validation set using the gini model

```{r}
wine_pred_gini <- predict(object = wine_tree, 
        newdata = wine_test, 
        type = "class")

str(wine_pred_gini)

```


# Compare classification error


```{r}

ce(actual = wine_test$quality, 
   predicted = wine_pred_gini)

```


```{r}

```


```{r}

```

<br>
<br>
